---
title: "AI2Analytics Quickstart (R)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AI2Analytics Quickstart (R)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

The smallest end-to-end example: synthetic HCP data → segmentation → cluster
profiles → diagnostic plot.

```{r}
library(ai2analytics)
```

## 1. Generate synthetic HCP data

5,000 fake providers with two prescribing-volume features. Bundled with the
package -- no downloads, no PII.

```{r}
tables <- generate_us_hcp(seed = 42L)
hcps   <- tables$hcp_reference
head(hcps)
```

## 2. Run the segmentation pipeline

```{r}
cfg <- SegmentationConfig(
  analysis_name   = "quickstart",
  col_entity_id   = "npi",
  feature_columns = c("IL_17_TRX_L12M", "IL_23_TRX_L12M"),
  n_segments      = 4
)

pipe <- SegmentationPipeline$new()
out  <- pipe$run(cfg, dataframes = list(entity_data = hcps))
```

## 3. Inspect the segments

```{r}
out$profiles
out$summary_stats[c("n_entities", "n_segments", "method", "silhouette_score")]
```

## 4. Diagnostic plot

```{r}
pipe$dashboard()
```

## What just happened

| Step | What it did |
|---|---|
| `generate_us_hcp()`              | Built a realistic HCP reference table (no PII). |
| `SegmentationConfig(...)`        | Declared which column is the entity ID and which are the features. |
| `SegmentationPipeline$new()$run(...)` | Normalised features, fitted KMeans, attached labels, computed silhouette, optionally wrote a CSV. |

Next: `01_one_template_three_regions.Rmd` shows the same pipeline reused
across US, EU, and BRIC schemas.
