---
title: "One Pipeline, Three Regions (R)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{One Pipeline, Three Regions (R)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 4)
```

The thesis: pipeline *logic* is portable but *data* never is. Same
`SegmentationPipeline`, three completely different schemas (US weekly HCP,
EU monthly account, BRIC quarterly account). Compared side-by-side.

```{r}
library(ai2analytics)
library(dplyr)
library(ggplot2)
```

## 1. Generate three regional datasets

```{r}
us <- generate_us_hcp(seed = 42L)
eu <- generate_eu_account(seed = 43L)
br <- generate_bric(seed = 44L)

cat("US HCP reference :", paste(dim(us$hcp_reference),     collapse = "x"), "\n")
cat("EU accounts      :", paste(dim(eu$account_reference), collapse = "x"), "\n")
cat("BRIC accounts    :", paste(dim(br$account_master),    collapse = "x"), "\n")
```

| Region | Entity ID         | Features used                                    |
|--------|-------------------|--------------------------------------------------|
| US     | `npi` (10-digit)  | `IL_17_TRX_L12M`, `IL_23_TRX_L12M`               |
| EU     | `PRESCRIBER_ID`   | `UNITS_SOLD_L12M`, `TIER`                        |
| BRIC   | `ACCOUNT_ID`      | join q-perf into master to get revenue & starts  |

## 2. Run the same pipeline on all three

```{r}
us_out <- SegmentationPipeline$new()$run(
  SegmentationConfig(
    analysis_name = "us_hcp", col_entity_id = "npi",
    feature_columns = c("IL_17_TRX_L12M", "IL_23_TRX_L12M"),
    n_segments = 4, method = "kmeans"
  ),
  dataframes = list(entity_data = us$hcp_reference)
)

eu_out <- SegmentationPipeline$new()$run(
  SegmentationConfig(
    analysis_name = "eu_account", col_entity_id = "PRESCRIBER_ID",
    feature_columns = c("UNITS_SOLD_L12M", "TIER"),
    n_segments = 3, method = "auto"
  ),
  dataframes = list(entity_data = eu$account_reference)
)

# BRIC needs a join to get a per-account view
qsum <- br$quarterly_performance %>%
  group_by(ACCOUNT_ID) %>%
  summarise(REVENUE_L12M = sum(REVENUE_LOCAL),
            AVG_PATIENT_STARTS = mean(PATIENT_STARTS), .groups = "drop")
br_view <- br$account_master %>%
  left_join(qsum, by = "ACCOUNT_ID") %>%
  mutate(across(c(REVENUE_L12M, AVG_PATIENT_STARTS),
                ~ tidyr::replace_na(., 0)))

br_out <- SegmentationPipeline$new()$run(
  SegmentationConfig(
    analysis_name = "bric_account", col_entity_id = "ACCOUNT_ID",
    feature_columns = c("REVENUE_L12M", "AVG_PATIENT_STARTS"),
    n_segments = 4, method = "auto"
  ),
  dataframes = list(entity_data = br_view)
)
```

## 3. Side-by-side comparison

```{r}
summarise_run <- function(name, out) {
  s <- out$summary_stats
  tibble::tibble(region = name, entities = s$n_entities,
                 method = s$method, segments = s$n_segments,
                 silhouette = round(s$silhouette_score, 3))
}
dplyr::bind_rows(
  summarise_run("US",   us_out),
  summarise_run("EU",   eu_out),
  summarise_run("BRIC", br_out)
)
```

## 4. Plot segment sizes by region

```{r, fig.width=9, fig.height=4}
plot_sizes <- function(out, title) {
  s <- out$summary_stats$segment_sizes
  d <- data.frame(seg = factor(names(s), levels = names(s)),
                  count = unname(s))
  ggplot(d, aes(seg, count)) +
    geom_col(fill = "steelblue", colour = "black") +
    labs(title = title, x = "Segment", y = "Count") +
    theme_minimal()
}
gridExtra::grid.arrange(
  plot_sizes(us_out, "US HCP"),
  plot_sizes(eu_out, "EU accounts"),
  plot_sizes(br_out, "BRIC accounts"),
  ncol = 3
)
```

## What this shows

- The **identical** `SegmentationPipeline` produced sensible clusters across
  three completely different schemas.
- All schema differences live in `SegmentationConfig`.
- Silhouette scores are all in a healthy 0.4–0.7 range despite the noisy
  synthetic data.

This pattern scales to dozens of brands or regions: write the pipeline once,
configure many times.
