---
title: "Market Mix Model: Where Should the Next Dollar Go? (R)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Market Mix Model: Where Should the Next Dollar Go? (R)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 4)
set.seed(42)
```

We fit `MarketMixPipeline` on synthetic weekly media + sales data, then use
the response curves to answer: **if I shift a dollar from Print to Digital,
what happens to incremental sales?**

```{r}
library(ai2analytics)
library(ggplot2)
```

## 1. Build a synthetic media + sales time-series

```{r}
n_weeks <- 156  # 3 years
dates   <- seq.Date(from = as.Date("2023-01-06"), by = "week", length.out = n_weeks)

trend       <- seq(100, 160, length.out = n_weeks)
seasonality <- 25 * sin(2 * pi * seq_len(n_weeks) / 52)
tv  <- pmin(250, rexp(n_weeks, rate = 1 / 60))
dig <- pmin(180, rexp(n_weeks, rate = 1 / 35))
prn <- pmin(120, rexp(n_weeks, rate = 1 / 25))

sales <- trend + seasonality +
  0.85 * log1p(tv)  * 12 +
  0.55 * log1p(dig) * 9  +
  0.25 * log1p(prn) * 5  +
  rnorm(n_weeks, 0, 8)

ts <- data.frame(
  WEEK_ENDING      = dates,
  SALES            = round(sales),
  TV_SPEND         = round(tv,  2),
  DIGITAL_SPEND    = round(dig, 2),
  PRINT_SPEND      = round(prn, 2),
  PRICE_INDEX      = round(100 + rnorm(n_weeks, 0, 2), 1),
  DISTRIBUTION_PCT = pmin(100, pmax(70, round(85 + rnorm(n_weeks, 0, 3), 1)))
)
head(ts)
```

## 2. Fit the model

```{r}
cfg <- MarketMixConfig(
  analysis_name      = "brand_x_mmm",
  col_date           = "WEEK_ENDING",
  col_response       = "SALES",
  media_columns      = c("TV_SPEND", "DIGITAL_SPEND", "PRINT_SPEND"),
  control_columns    = c("PRICE_INDEX", "DISTRIBUTION_PCT"),
  default_decay_rate = 0.5,
  default_saturation = "log",
  model_type         = "ridge",
  alpha              = 1.0,
  positive_coefficients = TRUE
)

pipe <- MarketMixPipeline$new()
out  <- pipe$run(cfg, dataframes = list(time_series = ts))
out$channel_summary
```

## 3. Visual model fit

```{r}
c <- out$contributions
ggplot(c, aes(x = seq_len(nrow(c)))) +
  geom_line(aes(y = actual), colour = "black") +
  geom_line(aes(y = predicted), colour = "red", linetype = "dashed") +
  labs(x = "Week index", y = "Sales", title = "Actual (black) vs Predicted (red)") +
  theme_minimal()
out$model_diagnostics
```

## 4. Response curves

```{r}
mean_spend <- colMeans(ts[, cfg$media_columns])
ggplot(out$response_curves, aes(spend_level, response, colour = channel)) +
  geom_line(linewidth = 1.1) +
  geom_vline(data = data.frame(channel = names(mean_spend), x = mean_spend),
             aes(xintercept = x, colour = channel),
             linetype = "dotted", show.legend = FALSE) +
  labs(x = "Per-period spend", y = "Modelled marginal response",
       title = "Response curves (dotted line = current avg spend)") +
  theme_minimal()
```

## 5. Budget what-if

A small grid search over reallocation moves at constant total spend.

```{r}
predicted_sales <- function(weekly_alloc, summary, decay = cfg$default_decay_rate) {
  ann <- 0
  for (ch in names(weekly_alloc)) {
    coef <- summary$coefficient[summary$channel == ch]
    ss   <- weekly_alloc[[ch]] / max(1e-9, 1 - decay)
    ann  <- ann + coef * log1p(ss) * 52
  }
  ann
}

baseline <- as.list(colMeans(ts[, cfg$media_columns]))
total    <- sum(unlist(baseline))
base_pred <- predicted_sales(baseline, out$channel_summary)

grid <- expand.grid(tv_pct = seq(20, 80, by = 5),
                    digital_pct = seq(10, 80, by = 5))
grid$print_pct <- 100 - grid$tv_pct - grid$digital_pct
grid <- subset(grid, print_pct >= 5 & print_pct <= 80)

grid$pred_ann_sales <- mapply(function(t, d, p) {
  predicted_sales(list(TV_SPEND = total * t / 100,
                       DIGITAL_SPEND = total * d / 100,
                       PRINT_SPEND = total * p / 100),
                  out$channel_summary)
}, grid$tv_pct, grid$digital_pct, grid$print_pct)

best <- head(grid[order(-grid$pred_ann_sales), ], 5)
cat(sprintf("Current mix : TV=%.0f%% Digital=%.0f%% Print=%.0f%%  -> %.0f\n",
            baseline$TV_SPEND / total * 100,
            baseline$DIGITAL_SPEND / total * 100,
            baseline$PRINT_SPEND / total * 100,
            base_pred))
print(best)
cat(sprintf("\nBest move adds ~%.0f sales/yr at the same total spend.\n",
            best$pred_ann_sales[1] - base_pred))
```

## What this shows

- A modeller who chose the wrong saturation function would get wrong marginal
  ROI numbers. We knew the DGP used `log1p`, so we picked `default_saturation =
  "log"` -- in real life, you'd backtest several.
- The reallocation grid uses **only** quantities the pipeline already returns
  (`channel_summary` for coefficients, `response_curves` for shape). No bespoke
  math.
- The pattern (fit MMM → simulate budget moves) is the actionable handoff to
  brand teams.
